---
title: "Publication narrative and figure gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Publication narrative and figure gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(popgenVCF)
```

This gallery is generated from deterministic synthetic canonical-result records. It demonstrates publication structure and interpretation boundaries; it is not production biological evidence and does not replace approval of the canonical real-data baseline.

## Maintained nine-family fixture

```{r fixture}
results <- list(
  pca = structure(list(variance = c(0.34, 0.19), n_samples = 8L, n_snps = 120L), class = "PopgenVCFPCAResult"),
  ibs = structure(list(method = "SNPRelate IBS"), class = "PopgenVCFIBSResult"),
  tree = structure(list(), class = "PopgenVCFTreeResult"),
  diversity = structure(list(), class = "PopgenVCFDiversityResult"),
  fst = structure(list(method = "Weir and Cockerham 1984"), class = "PopgenVCFFSTResult"),
  amova = structure(list(), class = "PopgenVCFAMOVAResult"),
  dapc = structure(list(parameters = list(n_pca = 12L, n_da = 2L)), class = "PopgenVCFDAPCResult"),
  ibd = structure(list(), class = "PopgenVCFIBDResult"),
  ancestry = structure(list(backend = "ADMIXTURE", selected_k = 3L), class = "PopgenVCFAncestryResult")
)
project <- new_popgenvcf_project(
  "Publication gallery fixture",
  results = results,
  project_id = "00000000-0000-0000-0000-000000000930"
)
inventory <- publication_narrative_inventory(project)
inventory[, c("kind", "state", "citation_keys")]
```

## Completeness evidence

```{r completeness}
publication_narrative_completeness(inventory)
```

Every canonical family has one narrative owner, a methods statement, a caption/legend, citation keys, and a supplementary summary. The same table records `skipped`, `unavailable`, `incomplete`, `failed`, and `diagnostic-only` states without inventing results.

## Interpretation boundaries

| Family | Gallery interpretation boundary |
|---|---|
| PCA | Ordination axes summarize genotype variation; they do not prove discrete populations or causality. |
| IBS/MDS | Sample similarity is not population-level FST. |
| Neighbour joining | Branch lengths follow the recorded distance and are not a time-calibrated phylogeny. |
| Diversity | Estimates depend on filtering, sample size, missingness, and population assignment. |
| FST | Population differentiation is estimator- and sampling-dependent. |
| AMOVA | Variance partitioning follows the supplied hierarchy and distance definition. |
| DAPC | Discrimination among supplied groups is not model-based ancestry. |
| IBD | Geographic association does not by itself identify a causal process. |
| Ancestry | Cluster labels are arbitrary and K selection requires replicate and fit evidence. |

## Transparent fallback states

```{r fallback}
fallback_project <- new_popgenvcf_project(
  "Publication fallback fixture",
  modules = list(
    pca = list(status = "skipped", reason = "No variants passed QC."),
    ancestry = list(status = "diagnostic-only", reason = "Exploratory backend comparison.")
  ),
  results = list(
    ancestry = structure(
      list(backend = "ADMIXTURE", selected_k = 2L, status = "diagnostic-only"),
      class = "PopgenVCFAncestryResult"
    )
  ),
  project_id = "00000000-0000-0000-0000-000000000931"
)
publication_narrative_inventory(fallback_project)[, c("kind", "state", "reason")]
```

## Publication companion outputs

`generate_publication_bundle()` writes the following maintained evidence in addition to methods, captions, bibliography, artifact lineage, and checksums:

- `manuscript/narrative-inventory.tsv`;
- `manuscript/narrative-completeness.tsv`;
- `supplementary/analysis-summaries.tsv`;
- `provenance/publication.json`.

These files allow reviewers and release automation to distinguish complete scientific narratives from transparent fallback states.
